Uniform formula lower bounds for testing the existence of and counting satisfiable assignments

### 2022-06-27 article/main.tex
Strange mistake in LP

\documentclass[a4paper, 11pt]{article}
% \usepackage{ae,lmodern}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[USenglish]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

@@

We will now show that applying the Speedup rule once and
using exhaustive search to remove the logarithmic-size quantifier
yields an improved Slowdown rule for quantified classes.
We have the following:
\begin{corollary}[Extended Slowdown rule]\label{cor:extended_slowdown}
	If $\SAT{}\in \NDL{c}$, then for every $d > 1$, we have:
	\[\ldots (Q~n^{a}) (\neg Q~n^b) \NDL{d}
	\subseteq \ldots (Q~n^{a}) \NDL{c\cdot\max(d/2, a, b)}.\]
\end{corollary}
\begin{proof}
	Without loss of generality, assume that $Q = \forall$, and therefore $\neg Q = \exists$.
	In that case, for all $q, r > 0$, we have $(\exists n^q) \NTIME[n^r] \subseteq \NTIME[n^{\max(q,r)}] $.
	Applying Corollary~\ref{cor:speedup} on the $\NDL{d}$ part and then using the 
	above observation leads to the following:
	\begin{flalign*}
	\ldots (\forall n^{a}) (\exists n^b) \NDL{d} 
		&\subseteq \ldots (\forall n^{a}) (\exists n^b) \NTIME[n^{d/2}]\\	
		&\subseteq \ldots (\forall n^{a}) \NTIME[n^{\max(d/2, d)}]	
	\end{flalign*}

@@

\subsection{Details of the model}
Given an integer $k$ and a real $c$, we define the following interger linear program $P_{c,k}$ as follows:
\begin{equation}
	todo
\end{equation}
\todoh{describe model}
\todoh{link to code?}

>>>>

\documentclass[a4paper, 11pt]{article}
% \usepackage{ae,lmodern}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[USenglish]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{todonotes}
\usepackage{enumitem}
% \usepackage{algorithm}
% \usepackage[noend]{algpseudocode}

@@

We will now show that applying the Speedup rule once and
using exhaustive search to remove the logarithmic-size quantifier
yields an improved Slowdown rule for quantified classes.
We have the following:
\begin{corollary}[Extended Slowdown rule]\label{cor:extended_slowdown}
	If $\SAT{}\in \NDL{c}$, then for every $d > 1$, we have:
	\[\ldots (Q~n^{a}) (\neg Q~n^b) \NDL{d}
	\subseteq \ldots (Q~n^{a}) \NDL{c\cdot\max(d/2, a, b)}.\]
\end{corollary}
\begin{proof}
	Without loss of generality, assume that $Q = \forall$, and therefore $\neg Q = \exists$.
	In that case, for all $q, r > 0$, we have $(\exists n^q) \NTIME[n^r] \subseteq \NTIME[n^{\max(q,r)}] $.
	Applying Corollary~\ref{cor:speedup} on the $\NDL{d}$ part and then using the 
	above observation leads to the following:
	\begin{flalign*}
	\ldots (\forall n^{a}) (\exists n^b) \NDL{d} 
		&\subseteq \ldots (\forall n^{a}) (\exists n^b) \NTIME[n^{d/2}]\\	
		&\subseteq \ldots (\forall n^{a}) \NTIME[n^{\max(d/2, b)}]	
	\end{flalign*}

@@

\subsection{Details of the model}
Given an integer $k \ge 3$ and a real $c$, we define the interger linear program $P_{c,k}$ as follows:
\[
\begin{array}{rrcll}
	(P_{k,c}) &\max& 1&&\\
	\text{Initial speedup constraints:}&&&&\\
	&d_1&=&d_0 - 2x_1&\\
	&a_{1, -1}&=&1&\\
	&a_{1, -2}&\ge&x_1&\\
	&su_1&=&1&\\
	\text{Speedup constraints:}&&&&\\
	\forall i \ge 1, su_i = 1\Rightarrow &d_i&=&d_{i-1} - 2x_i&\\
	\forall i \ge 1, su_i = 1\Rightarrow &a_{i, j}&\ge&a_{i-1, j+1}& j = 1,\ldots,k-1\\
	\forall i \ge 1, su_i = 1\Rightarrow &a_{i, -2}&\ge&x_i&\\
	\forall i \ge 1, su_i = 1\Rightarrow &a_{i, -1}&=&1&\\
	\text{Slowdown constraints:}&&&&\\
	\forall i \ge 1, su_i = 0\Rightarrow &d_i&\ge&c\cdot d_{i-1}/2&\\
	\forall i \ge 1, su_i = 0\Rightarrow &d_i&\ge&c\cdot a_{i-1, -1}&\\
	\forall i \ge 1, su_i = 0\Rightarrow &d_i&\ge&c\cdot a_{i-1, -2}&\\
	\forall i \ge 1, su_i = 0\Rightarrow &a_{i, j}&=&a_{i-1, j-1}& j = 2,\ldots,k\\
	\text{Well-formed proof constraints:}&&&&\\
	\forall i \ge 1, & 2\sum_{t=1}^i su_{t} &\ge& i&\\ 
					 & 2\sum_{i=1}^k su_{i} &=& k - 1&\\ 
	\text{Require contradiction:}&&&&\\
	& d_k &<&d_0&\\
	\text{Variables:}&&&&\\
	&d_i &\in& [0,+\infty), &i =0,\ldots, k\\
	&a_{i,j} &\in& [1,+\infty), &i,j =1,\ldots, k\\
	&x_i &\in& [0,+\infty), &i =1,\ldots, k\\
	&su_{i} &\in& \{0,1\}, &i =1,\ldots, k
\end{array}
\]
The variables have the following meaning:
\begin{enumerate}
	\item $d_i$: verifier is $\NDL{d_i}$ on line $i$,
	\item $a_{i,k+1-j}$: exponent of the $j$-th-to-last quantifier on line $i$. 
	We use
	$a_{i,-j}$ as a shorthand notation for $a_{i,k+1-j}$.
	\item $x_i$: if line $i$ is the result of a speedup, it is a speedup with parameter $x_i$,
	\item $su_{i}$: 1 if and only if line $i$ results in the application of a speedup, 0 otherwise.
\end{enumerate}
The constains of the type ``$su_i = b$ implies $C$'' can be encoded using the so-called ``Big M method''\footnote{See \url{https://en.wikipedia.org/wiki/Big\_M\_method}.}. 
The constaint $d_k < d_0$ can be approximately encoded as $d_k \le d_0 - \delta$ for some small $\delta > 0$.

The proof leads to a contradiction if and only if $d_k < d_0$, 
that is, if and only if the ILP is feasible.
Using binary search on $c$, we can quickly find an upper and lower bounds on $c_k^*$.
